---
title: "Dissertation_analysis_v1"
output: html_document
date: "2025-04-08"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Setup 

## Libraries

```{r}
library(tidyverse)
library(jsonlite)
library(lme4)
library(parameters)
library(MASS)
library(merDeriv)
library(ordinal)
library(patchwork)
library(sjPlot)
library(ggeffects)
library(marginaleffects)
library(interactions)
library(gtsummary)
library(performance)
library(HLMdiag)
select <- dplyr::select
```

## Reading in data

```{r}
df_raw <- read.csv("../data/exp_data/diss_data_prolific.csv") %>%
  mutate(
    date = format(as.Date(date, format = "%d/%m/%Y"), "%Y-%m-%d")
  )

# Subject level data
subject_raw1 <- df_raw %>%
  filter(date > as.Date("2025-03-23")) %>%
  rowwise() %>%
  do(data.frame(fromJSON(.$subject, flatten = T))) %>%
  ungroup()

# Condition information 
condition_df <- data.frame(
  condition = c("P", "Q", "R", "S"),
  reward = c("low", "low", "high", "high"),
  control = c("low", "high", "low", "high")
)

# Causal query information (answers)
causal_query_answers <- read.csv("causal_query_answers.csv")

# Trial level data
upi_data <- subject_raw1 %>%
  select(-c(cond_order, counter_balance_order, causal_query_structure, causal_query_relationship)) %>%
  distinct() %>%
  select(upi, condition) %>%
  distinct() %>%
  map_df(~rep(.x, each = 186))

trial_raw <- df_raw %>% 
  filter(date > as.Date("2025-03-23")) %>%
  rowwise() %>%
  do(data.frame(fromJSON(.$results, flatten = T))) %>%
  ungroup() %>%
  cbind(upi_data) %>%
  unite("round", c(graph, cb_group), sep = "_", remove = FALSE) %>% # False so that counterbalance can be tracked
  rename(
    reward_region = reward # Rename this to not confuse with reward prevalence column in condition_df
  ) %>%
  full_join(condition_df, by = "condition") %>%
  mutate(
    trial_count = as.integer(trial_count),
    trial_count = trial_count + 1, # Increase round count by one so that it starts from 1 instead of 0
    condition = factor(condition, # Add condition labels 
                       levels = c("P", "Q", "R", "S"), 
                       labels = c("Low Reward Low Control", "Low Reward High Control", 
                                  "High Reward Low Control", "High Reward High Control"))
  ) 
```

# Concentration checks and exclusion criterion

```{r}
# Lost focus on trials for more than 10 timesteps 
no_focus <- trial_raw %>%
  group_by(upi) %>%
  count(has_focus) %>%
  filter(has_focus == 0 & n > 10) 

# Took more than 40 minutes to complete the experiment 
subject_times_raw <- subject_raw1 %>%
  distinct(upi, start_time, end_time) %>%
  mutate(
    start = as.POSIXct(start_time, format="%Y-%m-%dT%H:%M:%OSZ", tz="UTC"),
    end = as.POSIXct(end_time, format="%Y-%m-%dT%H:%M:%OSZ", tz="UTC"),
    time = as.numeric(difftime(end, start, units = "mins"))
  ) 

long_time <- subject_times_raw %>%
  filter(time > 40) 

# Participants who concentrated
subject_concentration <- subject_raw1 %>%
  select(-c(cond_order, counter_balance_order, causal_query_structure, causal_query_relationship)) %>%
  distinct() %>%
  # Filter out participants who had 3 or more false attempts on the task, self-reported being distracted, lost focus (on another tab or application) during the experiment, and took longer than 40 minutes to complete the experiment
  filter(false_attempts < 3 & concentration == "yes" & upi != as.character(no_focus[1,1]) & !(upi %in% long_time$upi)) %>%
  select(upi)
```

# Tidying up data

## Explanation of dataframes 

Round_df: Rounds 1-6 level data 
trial_df0: Placeholder
Trial_df1: Within the rounds, data for individual timesteps (swapped for counterbalancing)
subject_df: Subject level data

```{r}
subject_raw2 <- subject_raw1 %>%
  # Excluding participants who did not concentrate
  inner_join(subject_concentration, by = "upi") %>%
  # Matching causal query to see if it is correct 
  mutate(
    causal_query_relationship = ifelse(causal_query_relationship == "-1", "neg", "pos"),
  ) %>%
  unite("round", c(cond_order, counter_balance_order), sep = "_", remove = TRUE) %>%
  unite("causal_query", c(causal_query_structure, causal_query_relationship), sep = "_", remove = TRUE) %>%
  mutate( 
    causal_identified = paste(round, causal_query) %in% 
                        paste(causal_query_answers$round, causal_query_answers$causal_query)
  ) %>%
  # Adding informative condition columns 
  full_join(condition_df, by = "condition") %>%
  # Tallying hope scale data
  select(-c(hope3, hope5, hope7, hope11)) %>%
  mutate(
    across(c(hope1:hope12, pre_control:post_difficulty, age, engagement), function(x) as.integer(x)),
    across(c(trackpad, reward, control, upi), function(x) as.factor(x)),
    hope_scale = rowSums(across(hope1:hope12))
  ) %>%
  # Removing redundant columns
  select(-c(proid, condition:sigma, false_attempts, hope1:hope12, strategy:text_bug))
```

## trial_df data

```{r}
trial_df0 <- trial_raw %>%
  inner_join(subject_concentration, by = "upi") %>%
  group_by(upi, trial_count, round, condition, reward, control) %>%
  summarise(
    interventions = max(cum_int),
    score = max(cum_trial_score) - max(cum_int),
    successful_reward = sum(reward_region) 
  ) %>%
  ungroup()
```

```{r}
trial_df1 <- trial_raw %>%
  inner_join(subject_concentration, by = "upi") %>%
  mutate(
    # Swap x_val and y_val
    temp_val = ifelse(cb_group == "B", x_val, x_val),
    x_val = ifelse(cb_group == "B", y_val, x_val),
    y_val = ifelse(cb_group == "B", temp_val, y_val),
    
    # Swap x_int and y_int
    temp_int = ifelse(cb_group == "B", x_int, x_int),
    x_int = ifelse(cb_group == "B", y_int, x_int),
    y_int = ifelse(cb_group == "B", temp_int, y_int),
    
    # Changing to 1s and 0s for reward_region
    reward_region = as.integer(reward_region),
    
    # Changing name labels for trial_count for plotting
    trial_count = factor(trial_count, 
                       levels = c("1", "2", "3", "4", "5", "6"), 
                       labels = c("Training R1", "Training R2", "Training R3", "Test R1", "Test R2", "Test R3"))
  ) %>%
  select(-c(x_held, y_held, z_int, cum_int:cum_trial_score, round, cb_group, has_focus, temp_val, temp_int))
```

## round_df data

```{r}
round_df1 <- trial_df0 %>%
  select(upi, round, trial_count, interventions, successful_reward, score, condition) %>%
  full_join(subject_raw2, by = c("upi", "round")) %>%
  mutate(
    score = score.x,
    final_score = score.y,
    phase = ifelse(trial_count < 4, "training", "test"),
    condition = fct_relevel(condition, "High Reward High Control", "Low Reward High Control", "High Reward Low Control", "Low Reward Low Control")
  ) %>%
  select(-c(score.x, score.y)) 
```

## subject_df data

```{r}
subject_df1 <- round_df1 %>%
  select(upi, condition, reward, control, pre_control:engagement, hope_scale, final_score, start_time, end_time) %>%
  distinct() %>%
  mutate(
    start = as.POSIXct(start_time, format="%Y-%m-%dT%H:%M:%OSZ", tz="UTC"),
    end = as.POSIXct(end_time, format="%Y-%m-%dT%H:%M:%OSZ", tz="UTC"),
    time = as.numeric(difftime(end, start, units = "mins")),
    age = as.integer(age),
    gender = as.factor(gender),
    condition = fct_relevel(condition, "High Reward High Control", "Low Reward High Control", "High Reward Low Control", "Low Reward Low Control")
  ) 
```

# Counterfactual analysis

Note: Noise ignored, which was a sigma of 4 in high control, 18 in low during the experiment 
Theta = [-1, 0, 1] for high control
Theta = [-0.25, 0, 0.25] for low control

```{r}
# Omega for conditions was standard at 0.8
omega = 0.8

counterfactual_0 <- trial_df1 %>%
  group_by(upi, trial_count) %>%
  mutate(
    reward_region_lead = lead(reward_region),
    z_lead = lead(z_val),
    intervention = x_int + y_int
  ) %>%
  mutate(
    previous_x = lag(x_val),
    previous_y = lag(y_val),
    previous_z = lag(z_val),
  ) %>%
  rename( # Make the column headings more clear given now we're dealing with the previous and next timestep
    next_z = z_lead,
    current_x = x_val,
    current_y = y_val,
    current_z = z_val,
    current_reward = reward_region,
    next_reward = reward_region_lead
  ) %>%
  filter(intervention == 1) %>%
  mutate(
    control = ifelse( # Changing low control conditions to high in control variable for test rounds
      trial_count %in% c("Training R1", "Training R2", "Training R3") & condition %in% c("Low Reward Low Control", "High Reward Low Control"),
      "low", "high")
  )

# Because timestep 30 is the last timesteps, all interventions are inconsequential 
counterfactual_inconsequential_30 <- counterfactual_0 %>%
  filter(step_counter == 30) %>%
  ungroup() %>%
  mutate(
    counterfactual = "Inconsequential"
  ) 

# Filter out timestep 30 that is in the other df 
counterfactual_1 <- counterfactual_0 %>% 
  filter(step_counter != 30)

# Calculating counterfactual X or Y values if participant had not intervened at the current timestep 
counterfactual_current <- counterfactual_1 %>%
  mutate(
    # Updating current x values based on causal graph and condition 
    current_x_updated = case_when(
      # If X had not been intervened on, retain current value as calculated in JS
      x_int == "0" ~ current_x, 
      
      # If high control (theta weights = 1 or -1)
      x_int == "1" & graph == "1" & control == "high" ~ omega * (-previous_x) + previous_x,
      x_int == "1" & graph == "2" & control == "high" ~ omega * (-previous_y - previous_x) + previous_x,
      x_int == "1" & graph == "3" & control == "high" ~ omega * (previous_y - previous_x) + previous_x,
      x_int == "1" & graph == "4" & control == "high" ~ omega * (-previous_y + previous_z - previous_x) + previous_x,
      x_int == "1" & graph == "5" & control == "high" ~ omega * (previous_z - previous_x) + previous_x,
      x_int == "1" & graph == "6" & control == "high" ~ omega * (-previous_y - previous_z - previous_x) + previous_x,
      
      # If low control (theta weights are 0.25 or -0.25, which is the expected value of a beta(1,3) distribution)
      x_int == "1" & graph == "1" & control == "low" ~ omega * (-previous_x) + previous_x,
      x_int == "1" & graph == "2" & control == "low" ~ omega * ((-0.25 * previous_y) - previous_x) + previous_x,
      x_int == "1" & graph == "3" & control == "low" ~ omega * ((0.25 * previous_y) - previous_x) + previous_x,
      x_int == "1" & graph == "4" & control == "low" ~ omega * ((-0.25 * previous_y) + (0.25 * previous_z) - previous_x) + previous_x,
      x_int == "1" & graph == "5" & control == "low" ~ omega * ((0.25 * previous_z) - previous_x) + previous_x,
      x_int == "1" & graph == "6" & control == "low" ~ omega * ((-0.25 * previous_y) + (-0.25 * previous_z) - previous_x) + previous_x,
    ),
    
    current_x_updated = round(current_x_updated),
    
    # Values are bounded between -100 and 100 
    current_x_updated = ifelse(current_x_updated > 100, 100, ifelse(current_x_updated < -100, -100, current_x_updated)),
    
    # Updating current y values based on causal graph and condition  
    current_y_updated = case_when(
      # If Y had not been intervened on, retain current value as calculated in JS
      y_int == "0" ~ current_y, 
      
      # If high control 
      y_int == "1" & graph == "1" & control == "high" ~ omega * (-0.25 * previous_x - previous_y) + previous_y,
      y_int == "1" & graph %in% c(2:6) & control == "high" ~ omega * (- previous_y) + previous_y,
      
      # If low control 
      y_int == "1" & graph == "1" & control == "low" ~ omega * ((-0.25 * previous_x) - previous_y) + previous_y,
      y_int == "1" & graph %in% c(2:6) & control == "low" ~ omega * (- previous_y) + previous_y,
      ),
    
    current_y_updated = round(current_y_updated),
    current_y_updated = ifelse(current_y_updated > 100, 100, ifelse(current_y_updated < -100, -100, current_y_updated)),
  )

# Calculating counterfactual Z values in next timestep if one hadn't intervened at current timestep 
counterfactual_next <- counterfactual_current %>%
  mutate(
    next_z_counterfactual = case_when(
      # If high control
      graph == "1" & control == "high" ~ omega * (-current_x_updated + current_y_updated - current_z) + current_z,  
      graph %in% c("2", "4") & control == "high" ~ omega * (-current_x_updated - current_z) + current_z,  
      graph %in% c("3", "5", "6") & control == "high" ~ omega * (-current_x_updated - current_y_updated - current_z) + current_z,
      
      # If low control
      graph == "1" & control == "low" ~ omega * ((-0.25 * current_x_updated) + (0.25 * current_y_updated) - current_z) + current_z,  
      graph %in% c("2", "4") & control == "low" ~ omega * ((-0.25 * current_x_updated) - current_z) + current_z,  
      graph %in% c("3", "5", "6") & control == "low" ~ omega * ((-0.25 * current_x_updated) + (-0.25 * current_y_updated) - current_z) + current_z,
      ),
    next_z_counterfactual = round(next_z_counterfactual),
    next_z_counterfactual = ifelse(next_z_counterfactual > 100, 100, ifelse(next_z_counterfactual < -100, -100, next_z_counterfactual)),
    
    counterfactual_reward = case_when(
      reward == "high" ~ ifelse(next_z_counterfactual >= 30 & next_z_counterfactual <= 70, 1, 0),
      reward == "low" ~ ifelse(next_z_counterfactual >= 40 & next_z_counterfactual <= 60, 1, 0)
      )
    )

counterfactual_df <- counterfactual_next %>%
  mutate(
    counterfactual = case_when(
        next_reward == 0 & counterfactual_reward == 0 ~ "Inconsequential",
        next_reward == 0 & counterfactual_reward == 1 ~ "Detrimental",
        next_reward == 1 & counterfactual_reward == 0 ~ "Necessary",
        next_reward == 1 & counterfactual_reward == 1 ~ "Redundant"
        )
  ) %>%
  rbind(counterfactual_inconsequential_30) # Adding back the inconsequential interventions at time = 30
```

## Proportion of types of interventions in training phase by condition 

```{r}
training_prop <- counterfactual_df %>% 
  filter(trial_count %in% c("Training R1", "Training R2", "Training R3")) %>%
  group_by(condition) %>%
  count(counterfactual) %>%
  mutate(
    proportion = n / sum(n),
    condition = fct_relevel(condition, "High Reward High Control", "Low Reward High Control", "High Reward Low Control", "Low Reward Low Control"),
    counterfactual = factor(counterfactual, levels = c("Redundant","Detrimental", "Necessary", "Inconsequential")),
    sum = sum(n)
    )

training_prop
```

```{r}
training_prop %>%
  select(condition, n) %>%
  group_by(condition) %>%
  mutate(
    sum = sum(n)
  )
```
  
```{r}
counterfactual_plot1 <- training_prop %>%
  ggplot(aes(x = condition, y = proportion, fill = counterfactual)) +
  geom_col(position = "stack", width = 0.75) +
  geom_text(aes(label = paste0(round(proportion, 3) * 100, "%")),
            size = ifelse(training_prop$proportion < 0.04, 2.5, 4), 
            position = position_stack(vjust = 0.5)) + 
            # color = ifelse(training_prop$intervention == "1", 'black', 'white')) +
  scale_fill_manual("counterfactual",
                    values = c("Detrimental" = "#E9C8B1",
                               "Redundant" = "#DBAD9C",
                               "Necessary" = "#F0DBC5",
                               "Inconsequential" = "#C2A9A1"),
                    name = "The intervention was:") +
  labs(x = NULL, y = "Proportion", title = NULL) +
  scale_y_continuous(labels = scales::percent) +
  scale_x_discrete(labels = c("Low Reward Low Control" = "Low Control\nSmall Reward Region\n(n = 1124)", "Low Reward High Control" = "High Control\nSmall Reward Region\n(n = 1454)", "High Reward Low Control" = "Low Control\nLarge Reward Region\n(n = 1512)", "High Reward High Control" = "High Control\nLarge Reward Region\n(n = 1242)")) +
  theme_light() +
  theme(
    legend.position = "top",
    axis.text.x = element_text(size = 12),   # X-axis text size
    axis.title.y = element_text(size = 15),  # Y-axis title size
    axis.text.y = element_text(size = 12),   # Y-axis text size
    legend.text = element_text(size = 12),
    legend.title = element_text(size = 12)
    )

counterfactual_plot1
```

## Proportion of types of interventions in testing phase by condition 

```{r}
test_prop <- counterfactual_df %>% 
  filter(trial_count %in% c("Test R1", "Test R2", "Test R3")) %>%
  group_by(condition) %>%
  count(counterfactual) %>%
  mutate(
    proportion = n / sum(n),
    condition = fct_relevel(condition, "High Reward High Control", "Low Reward High Control", "High Reward Low Control", "Low Reward Low Control"),
    counterfactual = factor(counterfactual, levels = c("Redundant","Detrimental", "Necessary", "Inconsequential")),
    sum = sum(n)
  )
test_prop
```

```{r}
counterfactual_plot2 <- test_prop %>%
  ggplot(aes(x = condition, y = proportion, fill = counterfactual)) +
  geom_col(position = "stack", width = 0.75) +
  geom_text(aes(label = paste0(round(proportion, 3) * 100, "%")),
            size = ifelse(test_prop$proportion < 0.04, 2.5, 4), 
            position = position_stack(vjust = 0.5)) + 
            # color = ifelse(test_prop$intervention == "1", 'black', 'white')) +
  scale_fill_manual("counterfactual",
                    values = c("Detrimental" = "#E9C8B1",
                               "Redundant" = "#DBAD9C",
                               "Necessary" = "#F0DBC5",
                               "Inconsequential" = "#C2A9A1"),
                    name = "The intervention was:") +
  labs(x = NULL, y = "Proportion", title = NULL) +
  scale_y_continuous(labels = scales::percent) +
  scale_x_discrete(labels = c("Low Reward Low Control" = "Low Control\nSmall Reward Region\n(n = 947)", "Low Reward High Control" = "High Control\nSmall Reward Region\n(n = 1476)", "High Reward Low Control" = "Low Control\nLarge Reward Region\n(n = 1298)", "High Reward High Control" = "High Control\nLarge Reward Region\n(n = 1340)")) +
  theme_light() +
  theme(
    legend.position = "top",
    axis.text.x = element_text(size = 12),   # X-axis text size
    axis.title.y = element_text(size = 15),  # Y-axis title size
    axis.text.y = element_text(size = 12),   # Y-axis text size
    legend.text = element_text(size = 12),
    legend.title = element_text(size = 12)
    )

counterfactual_plot2
```

```{r}
counterfactual_plot1 / counterfactual_plot2
```

## Redoing the graphs with raw counts instead of percentages 

```{r}
counterfactual_plot3 <- training_prop %>%
  ungroup() %>%
  complete(
    condition,
    counterfactual,
    fill = list(n = 0, proportion = 0)
  ) %>%
  ggplot(aes(x = condition, y = n, fill = counterfactual)) +
  geom_col(position = position_dodge(width = 0.8), width = 0.7) +
  geom_text(aes(label = n),
            size = 3.5,
            # size = ifelse(training_prop$n < 20, 2.5, 4), 
            position = position_dodge(width = 0.8), 
            vjust = -0.3) + 
  scale_fill_manual(
    values = c(
      "Detrimental" = "#E9C8B1",
      "Redundant" = "#DBAD9C",
      "Necessary" = "#F0DBC5",
      "Inconsequential" = "#C2A9A1"
    ),
    name = "The intervention was:"
  ) +
  labs(
    x = NULL,
    y = "Intervention Count",
    title = NULL
  ) +
  scale_x_discrete(labels = c(
    "Low Reward Low Control" = "Low Control\nSmall Reward Region",
    "Low Reward High Control" = "High Control\nSmall Reward Region",
    "High Reward Low Control" = "Low Control\nLarge Reward Region",
    "High Reward High Control" = "High Control\nLarge Reward Region"
  )) +
  scale_y_continuous(breaks = seq(0, 1250, by = 250)) +
  ylim(0, 1400) +
  theme_light() +
  theme(
    legend.position = "top",
    axis.text.x = element_text(size = 12),
    axis.title.y = element_text(size = 15),
    axis.text.y = element_text(size = 12),
    legend.text = element_text(size = 12),
    legend.title = element_text(size = 12)
  )

counterfactual_plot3
```

```{r}
counterfactual_plot4 <- test_prop %>%
  ggplot(aes(x = condition, y = n, fill = counterfactual)) +
  geom_col(position = position_dodge(width = 0.8), width = 0.7) +
  geom_text(aes(label = n),
            size = 3.5,
            # size = ifelse(training_prop$n < 20, 2.5, 4), 
            position = position_dodge(width = 0.8), 
            vjust = -0.3) + 
  scale_fill_manual(
    values = c(
      "Detrimental" = "#E9C8B1",
      "Redundant" = "#DBAD9C",
      "Necessary" = "#F0DBC5",
      "Inconsequential" = "#C2A9A1"
    ),
    name = "The intervention was:"
  ) +
  labs(
    x = NULL,
    y = "Intervention Count",
    title = NULL
  ) +
  scale_x_discrete(labels = c(
    "Low Reward Low Control" = "Low Control\nSmall Reward Region",
    "Low Reward High Control" = "High Control\nSmall Reward Region",
    "High Reward Low Control" = "Low Control\nLarge Reward Region",
    "High Reward High Control" = "High Control\nLarge Reward Region"
  )) +
  scale_y_continuous(breaks = seq(0, 1250, by = 250)) +
  ylim(0, 1400) +
  theme_light() +
  theme(
    legend.position = "top",
    axis.text.x = element_text(size = 12),
    axis.title.y = element_text(size = 15),
    axis.text.y = element_text(size = 12),
    legend.text = element_text(size = 12),
    legend.title = element_text(size = 12)
  )

counterfactual_plot4
```

### Combined plot just facetted 

```{r}
combined_prop <- counterfactual_df %>% 
  mutate(
    phase = ifelse(trial_count %in% c("Training R1", "Training R2", "Training R3"), "Training", "Test"),
    phase = fct_relevel(as.factor(phase), "Training")
  ) %>%
  group_by(phase, condition) %>%
  count(counterfactual) %>%
  mutate(
    proportion = n / sum(n),
    condition = fct_relevel(condition, "High Reward High Control", "Low Reward High Control", "High Reward Low Control", "Low Reward Low Control"),
    counterfactual = factor(counterfactual, levels = c("Redundant","Detrimental", "Necessary", "Inconsequential")),
    sum = sum(n)
    ) %>%
  ungroup() %>%
  complete(
    condition,
    counterfactual,
    phase, 
    fill = list(n = 0, proportion = 0)
  ) 

counterfactual_plot5 <- combined_prop %>%
  ggplot(aes(x = condition, y = n, fill = counterfactual)) +
  geom_col(position = position_dodge(width = 0.8), width = 0.7) +
  geom_text(aes(label = n),
            size = 3.5,
            # size = ifelse(training_prop$n < 20, 2.5, 4), 
            position = position_dodge(width = 0.8), 
            vjust = -0.3) + 
  scale_fill_manual(
    values = c(
      "Detrimental" = "#E9C8B1",
      "Redundant" = "#DBAD9C",
      "Necessary" = "#F0DBC5",
      "Inconsequential" = "#C2A9A1"
    ),
    name = "The intervention was:"
  ) +
  labs(
    x = NULL,
    y = "Intervention Count",
    title = NULL
  ) +
  scale_x_discrete(labels = c(
    "Low Reward Low Control" = "Low Control\nSmall Reward Region",
    "Low Reward High Control" = "High Control\nSmall Reward Region",
    "High Reward Low Control" = "Low Control\nLarge Reward Region",
    "High Reward High Control" = "High Control\nLarge Reward Region"
  )) +
  scale_y_continuous(breaks = seq(0, 1250, by = 250)) +
  ylim(0, 1400) +
  facet_wrap(~phase) +
  theme_light() +
  theme(
    strip.background = element_rect(fill = "#f5efd7", color = "black"),
    strip.text = element_text(color = "black", face = "bold", size = 12),
    legend.position = "top",
    axis.text.x = element_text(size = 8),
    axis.title.y = element_text(size = 15),
    axis.text.y = element_text(size = 12),
    legend.text = element_text(size = 12),
    legend.title = element_text(size = 12)
  )

counterfactual_plot5

```

```{r}
combined_prop %>%
  filter(phase == "Training") %>%
  select(condition, counterfactual, n)
```



```{r}
ggsave("../figures/analysis_figures/counterfactual_counts.png", counterfactual_plot5, height = 5, width = 10.5, dpi = 600)
```

# Normalised counts 

```{r}
counterfactual_plot6 <- combined_prop %>%
  full_join(count(subject_df1, condition, name = "condition_n")) %>%
  mutate(
    average_n = round(n / condition_n, digits = 1)
  ) %>%
  ggplot(aes(x = condition, y = average_n, fill = counterfactual)) +
  geom_col(position = position_dodge(width = 0.8), width = 0.7) +
  geom_text(aes(label = average_n),
            size = 3.5,
            # size = ifelse(training_prop$n < 20, 2.5, 4), 
            position = position_dodge(width = 0.8), 
            vjust = -0.3) + 
  scale_fill_manual(
    values = c(
      "Detrimental" = "#E9C8B1",
      "Redundant" = "#DBAD9C",
      "Necessary" = "#F0DBC5",
      "Inconsequential" = "#C2A9A1"
    ),
    name = "The intervention was:"
  ) +
  labs(
    x = NULL,
    y = "Intervention Count",
    title = NULL
  ) +
  scale_x_discrete(labels = c(
    "Low Reward Low Control" = "Low Control\nSmall Reward Region",
    "Low Reward High Control" = "High Control\nSmall Reward Region",
    "High Reward Low Control" = "Low Control\nLarge Reward Region",
    "High Reward High Control" = "High Control\nLarge Reward Region"
  )) +
  # scale_y_continuous(breaks = seq(0, 1250, by = 250)) +
  # ylim(0, 1400) +
  facet_wrap(~phase) +
  theme_light() +
  theme(
    strip.background = element_rect(fill = "#f5efd7", color = "black"),
    strip.text = element_text(color = "black", face = "bold", size = 12),
    legend.position = "top",
    axis.text.x = element_text(size = 8),
    axis.title.y = element_text(size = 15),
    axis.text.y = element_text(size = 12),
    legend.text = element_text(size = 12),
    legend.title = element_text(size = 12)
  )

counterfactual_plot6
```


```{r}
counterfactual_plot5 <- combined_prop %>%
  ggplot(aes(x = condition, y = n, fill = counterfactual)) +
  geom_col(position = position_dodge(width = 0.8), width = 0.7) +
  geom_text(aes(label = n),
            size = 3.5,
            # size = ifelse(training_prop$n < 20, 2.5, 4), 
            position = position_dodge(width = 0.8), 
            vjust = -0.3) + 
  scale_fill_manual(
    values = c(
      "Detrimental" = "#E9C8B1",
      "Redundant" = "#DBAD9C",
      "Necessary" = "#F0DBC5",
      "Inconsequential" = "#C2A9A1"
    ),
    name = "The intervention was:"
  ) +
  labs(
    x = NULL,
    y = "Intervention Count",
    title = NULL
  ) +
  scale_x_discrete(labels = c(
    "Low Reward Low Control" = "Low Control\nSmall Reward Region",
    "Low Reward High Control" = "High Control\nSmall Reward Region",
    "High Reward Low Control" = "Low Control\nLarge Reward Region",
    "High Reward High Control" = "High Control\nLarge Reward Region"
  )) +
  scale_y_continuous(breaks = seq(0, 1250, by = 250)) +
  ylim(0, 1400) +
  facet_wrap(~phase) +
  theme_light() +
  theme(
    strip.background = element_rect(fill = "#f5efd7", color = "black"),
    strip.text = element_text(color = "black", face = "bold", size = 12),
    legend.position = "top",
    axis.text.x = element_text(size = 8),
    axis.title.y = element_text(size = 15),
    axis.text.y = element_text(size = 12),
    legend.text = element_text(size = 12),
    legend.title = element_text(size = 12)
  )

counterfactual_plot5
```


## Counterfactual type counts by upi, phase 

```{r}
counterfactual_count <- counterfactual_df %>%
  mutate(
    phase = sub(" .*", "", trial_count)
    ) %>%
  ungroup() %>%
  count(upi, phase, counterfactual) %>%
  pivot_wider(
    names_from = counterfactual, 
    values_from = n
  ) %>%
  mutate(
    across(everything(), ~ replace(., is.na(.), 0))
    ) %>%
  # Adding missing row where participant didn't intervene at all in test phase 
  rbind(data.frame( 
    upi = "beulpvfoea",
    phase = "Test",
    Inconsequential = "0",
    Necessary = "0",
    Detrimental = "0",
    Redundant = "0")
    ) %>%
  mutate(
    across(c("Inconsequential":"Redundant"), as.integer)
  )
```

# Subjective Control (self-report/perceived controllability)

## Subjective Control Raw Data Plot

```{r}
subjective_control <- subject_df1 %>%
  rename(
    pre_pcontrol = pre_control,
    post_pcontrol = post_control 
  ) %>%
  pivot_longer(cols = c(pre_pcontrol, post_pcontrol, pre_difficulty, post_difficulty), 
               names_to = c("group", ".value"), 
               names_sep = "_") %>%
  mutate(
    phase = fct_recode(group,
                       "Test" = "post",
                       "Training" = "pre"),
    phase = fct_relevel(phase, "Training"),
    control = fct_recode(control,
                        "low_control" = "low",
                        "high_control" = "high"),
    control = fct_relevel(control, "high_control"),
    reward = fct_recode(reward,
                        "low_reward" = "low",
                        "high_reward" = "high"),
    reward = fct_relevel(reward, "high_reward"),
    pcontrol_ordered = factor(pcontrol, levels = 0:10, ordered = TRUE),
    upi = as.factor(upi)
  ) %>%
  full_join(counterfactual_count, by = c("upi", "phase")) 
```

## Subjective Control Model

### Ordinal regression model clm 

```{r}
subject_control_mdl1 <- clm(pcontrol_ordered ~ Detrimental + Inconsequential + Necessary + Redundant, data = subjective_control)

summary(subject_control_mdl1)
```

```{r}
exp(coef(subject_control_mdl1)[11])
exp(coef(subject_control_mdl1)[12])
exp(coef(subject_control_mdl1)[13])
exp(coef(subject_control_mdl1)[14])
```

### Linear mixed effects model

```{r}
summary(lm(pcontrol ~ Detrimental + Inconsequential + Necessary + Redundant, data = subjective_control))

subject_control_mdl2 <- lmer(pcontrol ~ Detrimental + Inconsequential + Necessary + Redundant + (1 | upi), data = subjective_control)

model_parameters(subject_control_mdl2, method = "kr")
```

### Mixed effects ordinal regression model clmm 

```{r}
subject_control_mdl3 <- clmm(pcontrol_ordered ~ Detrimental + Inconsequential + Necessary + Redundant  + (1 | upi), data = subjective_control, Hess = TRUE, nAGQ=10)

summary(subject_control_mdl3)
```
Estimates

```{r}
coef(subject_control_mdl3)[11:14] %>%
  round(2) %>%
  as.data.frame() %>%
  remove_rownames()
```

SEs

```{r}
round(summary(subject_control_mdl3)$coefficients[11:14,"Std. Error"], 2) %>%
  as.data.frame() %>%
  remove_rownames()
```

Odds ratios 

```{r}
round(exp(coef(subject_control_mdl3)[11:14]), 2) %>%
  as.data.frame() %>%
  remove_rownames()
```

```{r}
as.data.frame(round(exp(confint(subject_control_mdl3)[11:14,]), 2)) %>%
  remove_rownames()
```

Threshold coefficients 

```{r}
as.data.frame(round(summary(subject_control_mdl3)$coefficients[1:10,c("Estimate", "Std. Error")], 2))
```

```{r}
tab_model(subject_control_mdl3)
```

```{r}
tbl_regression(subject_control_mdl3)
```

### Comparing fit of the three models 

```{r}
AIC(subject_control_mdl1, subject_control_mdl2, subject_control_mdl3)
```

# Control attempts - Number of interventions 

```{r}
round_intervention <- round_df1 %>%
  filter(phase == "test") %>%
  mutate(
    upi = as.factor(upi)
  ) %>%
  full_join(counterfactual_count %>%
              filter(phase == "Training") %>%
              select(-phase), by = "upi")

round_intervention_mdl_data <- round_intervention %>%
  mutate(
    interventions_z = scale(interventions)[,1],
    age_z = scale(age)[,1],
    control = fct_relevel(control, "high"),
    reward = fct_relevel(reward, "high"),
  )
```

## Intervention models

```{r}
round_interventions_mdl1 <- lmer(interventions_z ~ Detrimental + Inconsequential + Necessary + Redundant + (1 | upi), data = round_intervention_mdl_data, control = lmerControl(optimizer = "bobyqa"))

model_parameters(round_interventions_mdl1, ci_method="kr")
```

```{r}
round_interventions_mdl2 <- lmer(interventions_z ~ (Detrimental + Inconsequential + Necessary + Redundant) * causal_identified  + (1 | upi), data = round_intervention_mdl_data, control = lmerControl(optimizer = "bobyqa"))

model_parameters(round_interventions_mdl2, ci_method="kr")
```

```{r}
round_interventions_mdl3 <- lmer(interventions_z ~ Detrimental + Inconsequential + Necessary + Redundant + age_z + trackpad + pre_control + causal_identified + hope_scale + (1 | upi), data = round_intervention_mdl_data, control = lmerControl(optimizer = "bobyqa"))

model_parameters(round_interventions_mdl3, ci_method="kr")
```

```{r}
round_interventions_mdl4 <- lmer(interventions_z ~ (Detrimental + Inconsequential + Necessary + Redundant) * reward * control + (1 | upi), data = round_intervention_mdl_data, control = lmerControl(optimizer = "bobyqa"))

model_parameters(round_interventions_mdl4, ci_method="kr")
```

```{r}
AIC(round_interventions_mdl1, round_interventions_mdl3, round_interventions_mdl4)
```

```{r}
anova(round_interventions_mdl1, round_interventions_mdl3)
```


```{r fig.height=10}
round_interventions_mdl1_asmp <- check_model(round_interventions_mdl1) %>%
  plot() %>%
  wrap_plots()


#Save as image
# ggsave("final_figures/assumptions/interventions_mdl.png", round_interventions_mdl2_asmp, height = 17, width = 12, dpi = 600)
```

Level 1 Influential points

```{r}
suppressWarnings(round_inf1 <- hlm_influence(round_interventions_mdl1, level = 1))
dotplot_diag(round_inf1$cooksd, cutoff = 1) + 
  theme_minimal() +
  labs(x = NULL, y = "Cooks Distance")
```

UPI cluster level influential points 

```{r}
suppressWarnings(round_inf2 <- hlm_influence(round_interventions_mdl1, level = "upi"))
dotplot_diag(round_inf2$cooksd, cutoff = 1, index=round_inf2$upi) + 
  theme_minimal() +
  labs(x = NULL, y = "Cooks Distance")
```


```{r}
tab_model(round_interventions_mdl1, p.val = "kr", show.df = TRUE, show.stat = TRUE, show.se = TRUE,
          dv.labels = c("Number of interventions (z-scored)"),
          title = "Regression table for Intervention Model") 
```

```{r}
tab_model(round_interventions_mdl3, p.val = "kr", show.df = TRUE, show.stat = TRUE, show.se = TRUE)
```

## Appendix: Time step of first intervention, summary statistics and model

```{r}
intervention_timestep1 <- trial_raw %>%
  inner_join(subject_concentration, by = "upi") %>%
  filter(trial_count > 3 & (x_int == 1 | y_int == 1)) %>%
  group_by(upi, trial_count) %>%
  summarise(
    timestep = min(step_counter)
  ) 

required_trials <- trial_raw %>%
  inner_join(subject_concentration, by = "upi") %>%
  filter(trial_count > 3 & (x_int == 1 | y_int == 1)) %>%
  distinct(upi, condition, reward, control)%>%
  crossing(trial_count = c(4:6))

intervention_timestep2 <- required_trials %>%
  left_join(intervention_timestep1, by = c("upi", "trial_count")) %>%
  mutate(
    timestep = ifelse(is.na(timestep), 0, timestep)
    ) %>%
  full_join(counterfactual_count %>%
              filter(phase == "Training") %>%
              select(-phase), by = "upi")
```


```{r}
intervention_timestep_mdl_data <- intervention_timestep2 %>%
  mutate(
    timestep_z = scale(timestep)[,1],
    control = fct_relevel(as.factor(control), "high"),
    reward = fct_relevel(as.factor(reward), "high"),
  )

intervention_timestep_mdl1 <- lmer(timestep_z ~ Detrimental + Inconsequential + Necessary + Redundant + (1 | upi), data = intervention_timestep_mdl_data, control = lmerControl(optimizer = "bobyqa"))

model_parameters(intervention_timestep_mdl1, ci_method="kr")
```

```{r}
tab_model(intervention_timestep_mdl1, p.val = "kr", show.df = TRUE, show.stat = TRUE, show.se = TRUE,
          dv.labels = c("Number of interventions (z-scored)"),
          # pred.labels = c("Intercept",
          #                 "Control: Low",
          #                  "Reward region: Small",
          #                 "Control * Reward"),
          title = "Regression table for lmer model") 
```

Performing more inconsequential interventions decreased the timestep of first intervention

# Success at task 

```{r}
reward_region <- trial_raw %>%
  inner_join(subject_concentration, by = "upi") %>%
  mutate( 
    reward_region = as.integer(reward_region),
    ) %>%
  filter(trial_count > 3 & step_counter != 0) %>%
  group_by(upi, trial_count, reward, control) %>%
  summarise(
    in_region = sum(reward_region)
  ) %>%
  full_join(counterfactual_count %>%
              filter(phase == "Training") %>%
              select(-phase), by = "upi") %>%
  ungroup() 
```

### Regressing rewarded timesteps on intervention types 

```{r}
reward_region_mdl_data <- reward_region %>%
  mutate(
    in_region_z = scale(in_region)[,1],
    control = fct_relevel(as.factor(control), "high"),
    reward = fct_relevel(as.factor(reward), "high"),
  )

reward_region_mdl1 <- lmer(in_region_z ~ Detrimental + Inconsequential + Necessary + Redundant + (1 | upi), data = reward_region_mdl_data, control = lmerControl(optimizer = "bobyqa"))

model_parameters(reward_region_mdl1, ci_method="kr")
```

```{r}
tab_model(reward_region_mdl1, p.val = "kr", show.df = TRUE, show.stat = TRUE, show.se = TRUE,
          dv.labels = c("Number of interventions (z-scored)"),
          # pred.labels = c("Intercept",
          #                 "Control: Low",
          #                  "Reward region: Small"),
          title = "Regression table for lmer model") 
```

Observing successful (necessary and redundant) interventions in training phase improves performance in test phase.

# Succesful interventions in test phase

```{r}
successful_intervention <- trial_df1 %>%
  group_by(upi, trial_count) %>%
  mutate(
    reward_region_lead = lead(reward_region),
    z_lead = lead(z_val),
    intervention = x_int + y_int,
    condition = fct_relevel(condition, "High Reward High Control", "Low Reward High Control", "High Reward Low Control", "Low Reward Low Control")
  ) %>%
  filter(step_counter != 0 
         # & step_counter != 30
         ) %>%
  ungroup() %>%
  mutate(
    reward_region_lead = ifelse(is.na(reward_region_lead), 0, reward_region_lead),
    z_lead = ifelse(is.na(z_lead), 0, z_lead)
  )

successful_intervention_prop <- successful_intervention %>%
  filter(str_detect(as.character(trial_count), "Test"), intervention == "1") %>%
  group_by(condition) %>%
  count(reward_region_lead) %>%
  mutate(
    proportion = n / sum(n),
    reward_region_lead = as.factor(reward_region_lead)
    )
```

```{r}
successful_intervention %>%
  filter(str_detect(as.character(trial_count), "Test"), intervention == "1") %>%
  group_by(condition) %>%
  count(reward_region_lead) %>%
  mutate(
    proportion = n / sum(n),
    )
```

```{r}
successful_intervention_plot <- successful_intervention_prop %>%
  ggplot(aes(x = condition, y = proportion, fill = reward_region_lead)) +
  geom_col(position = "stack", width = 0.75) +
  # geom_hline(yintercept = 0.25, linetype = "dashed", color = "black", size = 0.8, alpha = 0.4) +
  geom_text(aes(label = paste0(round(proportion, 3) * 100, "%")),
            size = 4, 
            position = "stack", 
            vjust = 2, 
            color = ifelse(successful_intervention_prop$reward_region_lead == "1", 'black', 'white')) +
  scale_fill_manual("reward_region_lead",
                    values = c("0" = "#ff6c3b",
                               "1" = "#7ad658"),
                    labels=c("0" = "Unsuccessful",
                             "1" = "Successful"),
                    name = NULL) +
  labs(x = NULL, y = "Proportion", title = NULL) +
  scale_y_continuous(labels = scales::percent) +
  scale_x_discrete(labels = c("Low Reward Low Control" = "Low Control\nSmall Reward Region\n(n = 947)", "Low Reward High Control" = "High Control\nSmall Reward Region\n(n = 1476)", "High Reward Low Control" = "Low Control\nLarge Reward Region\n(n = 1298)", "High Reward High Control" = "High Control\nLarge Reward Region\n(n = 1340)")) +
  theme_light() +
  theme(
    legend.position = "top",
    axis.text.x = element_text(size = 12),   # X-axis text size
    axis.title.y = element_text(size = 15),  # Y-axis title size
    axis.text.y = element_text(size = 12),   # Y-axis text size
    legend.text = element_text(size = 12)   # Legend text size
    )  

successful_intervention_plot
```


## Successful intervention model 

```{r}
successful_intervention_data <- successful_intervention %>%
  filter(intervention == 1) %>%
  mutate(
    control = fct_relevel(as.factor(control), "high"),
    reward = fct_relevel(as.factor(reward), "high"),
  ) %>%
  full_join(counterfactual_count %>%
              filter(phase == "Training") %>%
              select(-phase), by = "upi")

successful_intervention_mdl1 <- glmer(reward_region_lead ~ Detrimental + Inconsequential + Necessary + Redundant + (1 | upi), family = "binomial", data = successful_intervention_data)

summary(successful_intervention_mdl1)
```

Estimates

```{r}
fixef(successful_intervention_mdl1) %>%
  round(2) %>%
  as.data.frame() %>%
  remove_rownames()
```

SEs

```{r}
summary(successful_intervention_mdl1)$coefficients[,"Std. Error"] %>%
  round(2) %>%
  as.data.frame() %>%
  remove_rownames()
```

Odds ratios 

```{r}
fixef(successful_intervention_mdl1) %>%
  exp() %>%
  round(2) %>%
  as.data.frame() %>%
  remove_rownames()
```

```{r}
tab_model(successful_intervention_mdl1)
```

So experiencing inconsequential interventions in training rounds reduces the likelihood of producing successful interventions in test phase, whilst performing necessary and redundant interventions has the opposite effect. So performing successful interventions in the training phase carries over into the test phase. 

# Causal relationship identification

## Were participants able to identify a causal link? 

```{r}
counterfactual_round_count <- counterfactual_df %>%
  mutate(
    phase = sub(" .*", "", trial_count)
    ) %>%
  ungroup() %>%
  count(upi, phase, trial_count, counterfactual) %>%
  pivot_wider(
    names_from = counterfactual, 
    values_from = n
  ) %>%
  mutate(
    trial_count = case_when(
      trial_count == "Training R1" ~ 1,
      trial_count == "Training R2" ~ 2,
      trial_count == "Training R3" ~ 3,
      trial_count == "Test R1" ~ 4,
      trial_count == "Test R2" ~ 5,
      trial_count == "Test R3" ~ 6),
    trial_count = as.integer(trial_count)
    ) %>%
  complete(upi, trial_count = 1:6) %>%
  mutate(
    phase = ifelse(trial_count < 4, "training", "test"),
    across(everything(), ~ replace(., is.na(.), 0))
  )

causal_identification <- round_df1 %>%
  select(upi, round, phase, trial_count, interventions, reward, control, score, causal_identified) %>%
  full_join(counterfactual_round_count, by = c("upi", "trial_count", "phase"))
```

## Causal identification model based on counterfactual intervention 

```{r}
causal_identification_mdl1 <- glm(causal_identified ~ Detrimental + Inconsequential + Necessary + Redundant, family = "binomial", data = causal_identification)

summary(causal_identification_mdl1)
```

```{r}
causal_identification_mdl2 <- glmer(causal_identified ~ Detrimental + Inconsequential + Necessary + Redundant + (1 | upi), family = "binomial", data = causal_identification)

summary(causal_identification_mdl2)
```

# Task performance (scores in test rounds)

```{r}
performance_mdl1 <- lmer(score ~ Detrimental + Inconsequential + Necessary + Redundant + (1 | upi), data = round_intervention_mdl_data)

model_parameters(performance_mdl1, ci_method="kr")
```

```{r}
tab_model(performance_mdl1, p.val = "kr", show.df = TRUE, show.stat = TRUE, show.se = TRUE,
          dv.labels = c("Score"),
          # pred.labels = c("Intercept",
          #                 "Control: Low",
          #                  "Reward region: Small"),
          title = "Regression table for lmer model") 
```




