---
title: "Simulation Analysis"
output: html_document
date: "2025-09-23"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(ggdist)
library(ggrain)
library(jsonlite)
```

# Reading in simulation data for random interventions and combining 

```{r}
df1 <- read.csv("../data/control_strategy_simulated_data/random_simulation_cond1.csv")
df2 <- read.csv("../data/control_strategy_simulated_data/random_simulation_cond2.csv")
df3 <- read.csv("../data/control_strategy_simulated_data/random_simulation_cond3.csv")
df4 <- read.csv("../data/control_strategy_simulated_data/random_simulation_cond4.csv")

condition_df1 <- data.frame(
  condition = c("1", "2", "3", "4"),
  reward = c("Small Reward Region", "Small Reward Region", "Large Reward Region", "Large Reward Region"),
  control = c("Low Control", "High Control", "Low Control", "High Control")
  ) %>%
  mutate(
    condition = as.integer(condition)
  )

random_df <- rbind(df1, df2, df3, df4) %>%
  full_join(condition_df1, by = "condition") %>%
  mutate(
    type = "Random interventions",
    trial_score = total_score
  ) %>%
  select(-total_score)
  

rm(df1, df2, df3, df4)
```

# Reading in simulation data for no interventions and combining 

```{r}
df1 <- read.csv("../data/control_strategy_simulated_data/simulation_cond1.csv")
df2 <- read.csv("../data/control_strategy_simulated_data/simulation_cond2.csv")
df3 <- read.csv("../data/control_strategy_simulated_data/simulation_cond3.csv")
df4 <- read.csv("../data/control_strategy_simulated_data/simulation_cond4.csv")

no_df <- rbind(df1, df2, df3, df4) %>%
  full_join(condition_df1, by = "condition") %>%
  mutate(
    type = "No interventions"
  ) %>%
  select(-total_score)

rm(df1, df2, df3, df4)
```

# Proportion scoring by timesteps 

## Random simulation



# Distribution of scores in random control models by condition 

```{r}
random_df %>% 
  ggplot(aes(x = trial_score)) +
  geom_bar(fill = "#029cd4", color = "black") +
  scale_x_continuous(breaks = seq(-30, 50, by = 10)) +
  facet_grid(reward~control) +
  labs(
    title = "Change in scores for simulated trials with random interventions each timestep",
  # (where interventions on X or Y across all values are equally likely)
    x = "Points Scored",
    y = NULL
  ) +
  theme_light() +
  theme(
    strip.background = element_rect(fill = "#f5efd7", color = "black"),
    strip.text = element_text(color = "black", face = "bold", size = 15),
    legend.position = "top",
    # plot.background = element_rect(fill = "transparent", color = NA)
    axis.title.x = element_text(size = 15),  # X-axis title size
    axis.text.x = element_text(size = 12),   # X-axis text size
    axis.title.y = element_text(size = 15),  # Y-axis title size
    axis.text.y = element_text(size = 12),   # Y-axis text size
    legend.text = element_text(size = 12)   # Legend text size
    )
```

```{r}
random_df %>% 
  filter(control == "High Control") %>%
  ggplot(aes(x = trial_score)) +
  geom_bar(fill = "#029cd4", color = "black") +
  scale_x_continuous(breaks = seq(-30, 50, by = 10)) +
  facet_wrap(~graph) +
  labs(
    title = "Change in scores for simulated trials with random interventions each timestep",
  # (where interventions on X or Y across all values are equally likely)
    x = "Points Scored",
    y = NULL
  ) +
  theme_light() +
  theme(
    strip.background = element_rect(fill = "#f5efd7", color = "black"),
    strip.text = element_text(color = "black", face = "bold", size = 15),
    legend.position = "top",
    # plot.background = element_rect(fill = "transparent", color = NA)
    axis.title.x = element_text(size = 15),  # X-axis title size
    axis.text.x = element_text(size = 12),   # X-axis text size
    axis.title.y = element_text(size = 15),  # Y-axis title size
    axis.text.y = element_text(size = 12),   # Y-axis text size
    legend.text = element_text(size = 12)   # Legend text size
    )
```

# Distribution of scores in no control models by condition 

```{r}
no_df %>% 
  ggplot(aes(x = trial_score)) +
  geom_bar(fill = "#029cd4", color = "black") +
  # scale_x_continuous(breaks = seq(-30, 50, by = 10)) +
  facet_grid(reward~control) +
  labs(
    title = "Change in scores for simulated trials with no interventions performed",
  # (where interventions on X or Y across all values are equally likely)
    x = "Points Scored",
    y = NULL
  ) +
  theme_light() +
  theme(
    strip.background = element_rect(fill = "#f5efd7", color = "black"),
    strip.text = element_text(color = "black", face = "bold", size = 15),
    legend.position = "top",
    # plot.background = element_rect(fill = "transparent", color = NA)
    axis.title.x = element_text(size = 15),  # X-axis title size
    axis.text.x = element_text(size = 12),   # X-axis text size
    axis.title.y = element_text(size = 15),  # Y-axis title size
    axis.text.y = element_text(size = 12),   # Y-axis text size
    legend.text = element_text(size = 12)   # Legend text size
    )
```

```{r}
no_df %>% 
  filter(control == "High Control") %>%
  ggplot(aes(x = trial_score)) +
  geom_bar(fill = "#029cd4", color = "black") +
  # scale_x_continuous(breaks = seq(-30, 50, by = 10)) +
  facet_wrap(~graph) +
  labs(
    title = "Change in scores for simulated trials with no interventions performed",
  # (where interventions on X or Y across all values are equally likely)
    x = "Points Scored",
    y = NULL
  ) +
  theme_light() +
  theme(
    strip.background = element_rect(fill = "#f5efd7", color = "black"),
    strip.text = element_text(color = "black", face = "bold", size = 15),
    legend.position = "top",
    # plot.background = element_rect(fill = "transparent", color = NA)
    axis.title.x = element_text(size = 15),  # X-axis title size
    axis.text.x = element_text(size = 12),   # X-axis text size
    axis.title.y = element_text(size = 15),  # Y-axis title size
    axis.text.y = element_text(size = 12),   # Y-axis text size
    legend.text = element_text(size = 12)   # Legend text size
    )
```

# Comparing to actual participant performance

## Data cleaning of participant scores

```{r}
df_raw <- read.csv("../data/exp_data/diss_data_prolific.csv") %>%
  mutate(
    date = format(as.Date(date, format = "%d/%m/%Y"), "%Y-%m-%d")
  )

# Subject level data
subject_raw1 <- df_raw %>%
  filter(date > as.Date("2025-03-23")) %>%
  rowwise() %>%
  do(data.frame(fromJSON(.$subject, flatten = T))) %>%
  ungroup()

# Condition information 
condition_df2 <- data.frame(
  condition = c("P", "Q", "R", "S"),
  reward = c("Small Reward Region", "Small Reward Region", "Large Reward Region", "Large Reward Region"),
  control = c("Low Control", "High Control", "Low Control", "High Control")
)

# Causal query information (answers)
causal_query_answers <- read.csv("causal_query_answers.csv")

# Trial level data
upi_data <- subject_raw1 %>%
  select(-c(cond_order, counter_balance_order, causal_query_structure, causal_query_relationship)) %>%
  distinct() %>%
  select(upi, condition) %>%
  distinct() %>%
  map_df(~rep(.x, each = 186))

trial_raw <- df_raw %>% 
  filter(date > as.Date("2025-03-23")) %>%
  rowwise() %>%
  do(data.frame(fromJSON(.$results, flatten = T))) %>%
  ungroup() %>%
  cbind(upi_data) %>%
  unite("round", c(graph, cb_group), sep = "_", remove = FALSE) %>% # False so that counterbalance can be tracked
  rename(
    reward_region = reward # Rename this to not confuse with reward prevalence column in condition_df
  ) %>%
  full_join(condition_df2, by = "condition") %>%
  mutate(
    trial_count = as.integer(trial_count),
    trial_count = trial_count + 1, # Increase round count by one so that it starts from 1 instead of 0
    condition = factor(condition, # Add condition labels 
                       levels = c("P", "Q", "R", "S"), 
                       labels = c("Low Reward Low Control", "Low Reward High Control", 
                                  "High Reward Low Control", "High Reward High Control"))
  ) 

# Lost focus on trials for more than 10 timesteps 
no_focus <- trial_raw %>%
  group_by(upi) %>%
  count(has_focus) %>%
  filter(has_focus == 0 & n > 10) 

# Took more than 40 minutes to complete the experiment 
subject_times_raw <- subject_raw1 %>%
  distinct(upi, start_time, end_time) %>%
  mutate(
    start = as.POSIXct(start_time, format="%Y-%m-%dT%H:%M:%OSZ", tz="UTC"),
    end = as.POSIXct(end_time, format="%Y-%m-%dT%H:%M:%OSZ", tz="UTC"),
    time = as.numeric(difftime(end, start, units = "mins"))
  ) 

long_time <- subject_times_raw %>%
  filter(time > 40) 
  


# Participants who concentrated
subject_concentration <- subject_raw1 %>%
  select(-c(cond_order, counter_balance_order, causal_query_structure, causal_query_relationship)) %>%
  distinct() %>%
  # Filter out participants who had 3 or more false attempts on the task, self-reported being distracted, lost focus (on another tab or application) during the experiment, and took longer than 40 minutes to complete the experiment
  filter(false_attempts < 3 & concentration == "yes" & upi != as.character(no_focus[1,1]) & !(upi %in% long_time$upi)) %>%
  select(upi)

trial_df <- trial_raw %>%
  inner_join(subject_concentration, by = "upi") %>%
  group_by(upi, trial_count, round, condition, reward, control) %>%
  summarise(
    interventions = max(cum_int),
    score = max(cum_trial_score) - max(cum_int),
    successful_reward = sum(reward_region) 
  ) %>%
  ungroup() %>%
  rename(
    trial_score = score
  ) %>%
  mutate(
    type = "Experiment results"
  )
  
rm(df_raw, subject_raw1, condition_df2, causal_query_answers, upi_data, trial_raw, no_focus, subject_times_raw, long_time, subject_raw1, subject_concentration)
```

```{r}
trial_df %>%
  mutate(
    phase = ifelse(trial_count %in% 1:3, "training", "test")
    ) %>%
  group_by(phase, condition) %>%
  summarise(
    mean_value = mean(trial_score, na.rm = TRUE),
    sd_value = sd(trial_score, na.rm = TRUE),
    n = n()
  )
```



## Participant scoring distributions by condition

```{r}
trial_df %>% 
  filter(trial_count < 4) %>%
  ggplot(aes(x = trial_score)) +
  geom_density(fill = "#029cd4", color = "black") +
  # scale_x_continuous(breaks = seq(-30, 50, by = 10)) +
  facet_grid(reward~control) +
  labs(
    title = "Actual change in score (training rounds)",
  # (where interventions on X or Y across all values are equally likely)
    x = "Points Scored",
    y = NULL
  ) +
  theme_light() +
  theme(
    strip.background = element_rect(fill = "#f5efd7", color = "black"),
    strip.text = element_text(color = "black", face = "bold", size = 15),
    legend.position = "top",
    # plot.background = element_rect(fill = "transparent", color = NA)
    axis.title.x = element_text(size = 15),  # X-axis title size
    axis.text.x = element_text(size = 12),   # X-axis text size
    axis.title.y = element_text(size = 15),  # Y-axis title size
    axis.text.y = element_text(size = 12),   # Y-axis text size
    legend.text = element_text(size = 12)   # Legend text size
    )
```

```{r}
trial_df %>% 
  filter(trial_count > 3) %>%
  ggplot(aes(x = trial_score)) +
  geom_density(fill = "#029cd4", color = "black") +
  # scale_x_continuous(breaks = seq(-30, 50, by = 10)) +
  facet_grid(reward~control) +
  labs(
    title = "Actual change in score (test rounds)",
  # (where interventions on X or Y across all values are equally likely)
    x = "Points Scored",
    y = NULL
  ) +
  theme_light() +
  theme(
    strip.background = element_rect(fill = "#f5efd7", color = "black"),
    strip.text = element_text(color = "black", face = "bold", size = 15),
    legend.position = "top",
    # plot.background = element_rect(fill = "transparent", color = NA)
    axis.title.x = element_text(size = 15),  # X-axis title size
    axis.text.x = element_text(size = 12),   # X-axis text size
    axis.title.y = element_text(size = 15),  # Y-axis title size
    axis.text.y = element_text(size = 12),   # Y-axis text size
    legend.text = element_text(size = 12)   # Legend text size
    )
```

# Combining graphs together to compare

```{r}
rbind(
  select(subset(trial_df, trial_count < 4), trial_score, type, reward, control),
  # select(no_df, trial_score, type, reward, control),
  select(random_df, trial_score, type, reward, control)
) %>%
ggplot(aes(x = trial_score, fill = type)) +
  
  geom_density(adjust = 3.5, alpha = 0.5, color = "black") +
  
  # geom_density(data = no_df1, aes(x = trial_score), adjust = 3, fill = "#C2A9A1", alpha = 0.5, color = "black") +
  
  # geom_density(data = subset(trial_df, trial_count < 4), aes(x = trial_score), fill = "#029cd4", alpha = 0.5, color = "black") +
  
  scale_y_continuous(
    limits = c(0, 0.1),
    breaks = seq(0, 0.1, by = 0.02)
    ) +
  
  scale_x_continuous(breaks = seq(-30, 120, by = 30)) +
  
  facet_grid(reward~control) +
  labs(
    x = "Points Scored",
    y = NULL
  ) +
  theme_light() +
  theme(
    strip.background = element_rect(fill = "white", color = "black"),
    strip.text = element_text(color = "black", face = "bold", size = 15),
    legend.position = "top",
    # plot.background = element_rect(fill = "transparent", color = NA)
    axis.title.x = element_text(size = 15),  # X-axis title size
    axis.text.x = element_text(size = 12),   # X-axis text size
    axis.title.y = element_text(size = 15),  # Y-axis title size
    axis.text.y = element_text(size = 12),   # Y-axis text size
    legend.text = element_text(size = 12),   # Legend text size
    legend.title = element_blank()
    )
```

# Combining graphs together to compare (only simulations )

```{r}
plot1 <- rbind(
  select(no_df, trial_score, type, reward, control),
  select(random_df, trial_score, type, reward, control)
) %>%
ggplot(aes(x = trial_score, fill = type)) +
  
  geom_bar(adjust = 3.5, alpha = 0.5, color = "black", position = "identity") +
  
  # geom_density(data = no_df1, aes(x = trial_score), adjust = 3, fill = "#C2A9A1", alpha = 0.5, color = "black") +
  
  # geom_density(data = subset(trial_df, trial_count < 4), aes(x = trial_score), fill = "#029cd4", alpha = 0.5, color = "black") +
  
  # scale_y_continuous(
  #   limits = c(0, 0.1),
  #   breaks = seq(0, 0.1, by = 0.02)
  #   ) +
  
  scale_x_continuous(breaks = seq(-20, 120, by = 20)) +
  
  scale_fill_manual(
    name = NULL,
    values = c("No interventions" = "#029cd4", "Random interventions" = "#d9bc00")
  ) +
  
  facet_grid(reward~control) +
  labs(
    x = "Points Scored",
    y = NULL
  ) +
  theme_light() +
  theme(
    strip.background = element_rect(fill = "#f5efd7", color = "black"),
    strip.text = element_text(color = "black", face = "bold", size = 15),
    legend.position = "top",
    # plot.background = element_rect(fill = "transparent", color = NA)
    axis.title.x = element_text(size = 15),  # X-axis title size
    axis.text.x = element_text(size = 12),   # X-axis text size
    axis.title.y = element_text(size = 15),  # Y-axis title size
    axis.text.y = element_text(size = 12),   # Y-axis text size
    legend.text = element_text(size = 12),   # Legend text size
    legend.title = element_blank()
    )

plot1
```

```{r}
rbind(
  select(no_df, trial_score, type, reward, control),
  select(random_df, trial_score, type, reward, control)
) %>%
  group_by(type, reward, control) %>%
  summarise(
    mean = mean(trial_score),
    sd = sd(trial_score)
  )
```




## Raincloud plot

```{r}
rbind(
  select(subset(trial_df, trial_count < 4), trial_score, type, reward, control),
  select(no_df, trial_score, type, reward, control),
  select(random_df, trial_score, type, reward, control)
) %>%
ggplot(aes(x = type, y = trial_score, fill = type)) +
  
  geom_rain(alpha = .5) +
  
  # stat_halfeye(
  #   # adjust bandwidth
  #   adjust = 0.5,
  #   # move to the right
  #   justification = -0.2,
  #   # remove the slub interval
  #   .width = 0,
  #   point_colour = NA
  # ) +
  # 
  # geom_boxplot(
  #   width = 0.12,
  #   # removing outliers
  #   outlier.color = NA,
  #   alpha = 0.5
  # ) +
  
  # scale_y_continuous(
  #   limits = c(0, 0.1),
  #   breaks = seq(0, 0.1, by = 0.02)
  #   ) +
  
  # scale_x_continuous(breaks = seq(-30, 120, by = 30)) +
  
  facet_grid(reward~control) +
  labs(
    title = "Comparison of score distributions between experimental results with no or random control (training rounds)",
  # (where interventions on X or Y across all values are equally likely)
    x = "Points Scored",
    y = NULL
  ) +
  theme_light() +
  theme(
    strip.background = element_rect(fill = "#f5efd7", color = "black"),
    strip.text = element_text(color = "black", face = "bold", size = 15),
    legend.position = "top",
    # plot.background = element_rect(fill = "transparent", color = NA)
    axis.title.x = element_text(size = 15),  # X-axis title size
    axis.text.x = element_text(size = 12),   # X-axis text size
    axis.title.y = element_text(size = 15),  # Y-axis title size
    axis.text.y = element_text(size = 12),   # Y-axis text size
    legend.text = element_text(size = 12),   # Legend text size
    legend.title = element_blank()
    ) +
  scale_fill_brewer(palette = 'Dark2')
```

TODO: tidy up raincloud plot 

# Comparing test round results with regained control

```{r}
random_df2 <- random_df %>%
  filter(control == "High Control") %>%
  mutate(
    control = "Low Control"
  ) %>%
  rbind(subset(random_df, control == "High Control"))

no_df2 <- no_df %>%
  filter(control == "High Control") %>%
  mutate(
    control = "Low Control"
  ) %>%
  rbind(subset(no_df, control == "High Control"))

simulated_scores <- rbind(
  # select(no_df2, trial_score, type, reward, control),
  select(random_df2, trial_score, type, reward, control),
  select(subset(trial_df, trial_count > 3), trial_score, type, reward, control)
  ) 

ggplot() +
  geom_density(data = simulated_scores, aes(x = trial_score, fill = type), adjust = 1, alpha = 0.5, color = "black") +

  scale_x_continuous(breaks = seq(-40, 120, by = 20)) +
  
  facet_grid(reward~control) +
  labs(
    x = "Points Scored",
    y = NULL
  ) +
  theme_light() +
  theme(
    strip.background = element_rect(fill = "white", color = "black"),
    strip.text = element_text(color = "black", face = "bold", size = 15),
    legend.position = "top",
    # plot.background = element_rect(fill = "transparent", color = NA)
    axis.title.x = element_text(size = 15),  # X-axis title size
    axis.text.x = element_text(size = 12),   # X-axis text size
    axis.title.y = element_text(size = 15),  # Y-axis title size
    axis.text.y = element_text(size = 12),   # Y-axis text size
    legend.text = element_text(size = 12),   # Legend text size
    legend.title = element_blank()
    )
```

```{r}
simulated_scores %>%
  group_by(type, reward, control) %>%
  summarise(
    mean = mean(trial_score),
    sd = sd(trial_score),
    median = median(trial_score)
  )
```




```{r}
trial_df %>%
  filter(control == "Low Control", reward == "Large Reward Region", trial_count > 3) %>%
  pull(trial_score) %>%
  var()

no_df %>%
  filter(control == "Low Control", reward == "Large Reward Region") %>%
  pull(trial_score) %>%
  var()
```


# Using t-tests to compare participant performance to random and no control models 

Welsh test examples comparing high control and large reward region conditions to random and no control models in the training phase. 

```{r}
t.test(
  trial_df %>%
    filter(control == "High Control", reward == "Large Reward Region", trial_count < 4) %>%
    pull(trial_score),
  
  no_df %>%
    filter(control == "High Control", reward == "Large Reward Region") %>%
    pull(trial_score),
  
  var.equal = FALSE
)
```

```{r}
t.test(
  trial_df %>%
    filter(control == "High Control", reward == "Large Reward Region", trial_count < 4) %>%
    pull(trial_score),
  
  random_df %>%
    filter(control == "High Control", reward == "Large Reward Region") %>%
    pull(trial_score),
  
  var.equal = FALSE
)
```

Now repeat that process for all condition X simulation X phase combinations. Test phase only has comparisons to high control simulations for both large and small reward regions. 

```{r}
condition_df2 <- condition_df1 %>%
  crossing(simulation = c("no", "random")) %>%
  crossing(phase = c("training", "test"))

results <- vector("list", nrow(condition_df2))  # empty list to store results

for (i in 1:nrow(condition_df2)) {
  row <- condition_df2[i, ]   # extract the i-th row as a 1-row dataframe
  
  # Access values like this:
  control1 <- row$control
  reward1 <- row$reward
  simulation <- row$simulation
  phase <- row$phase
  
  if (phase == "training") {
    
    df1 <- trial_df %>%
        filter(control == control1, reward == reward1, trial_count < 4) %>%
        pull(trial_score)
    
    df2 <- get(paste0(simulation, "_df")) %>%  # Merge simulation with _df and call the subsequent object 
        filter(control == control1, reward == reward1) %>%
        pull(trial_score)
    
  } else {
    
    df1 <- trial_df %>%
      filter(control == control1, reward == reward1, trial_count > 3) %>%
      pull(trial_score)
      
    df2 <- get(paste0(simulation, "_df2")) %>%  # Merge simulation with _df2 and call the subsequent object 
      filter(control == control1, reward == reward1) %>%
      pull(trial_score)
  }
  
  test_res <- t.test(df1, df2, var.equal = FALSE, alternative = "two.sided")
  
  # store key results in a list
  results[[i]] <- data.frame(
    t_stat = test_res$statistic,
    p_value = test_res$p.value,
    df = test_res$parameter,
    mean_x = test_res$estimate[1],
    sd_x = sd(df1),
    # var_x = var(df1),
    mean_y = test_res$estimate[2],
    sd_y = sd(df2)
    # var_y = var(df2)
  )
}

t_test_results <- bind_cols(condition_df2, bind_rows(results)) %>%
  mutate(
    p_value = ifelse(p_value < 0.001,
                      "< .001",
                      sprintf("%.3f", p_value))
  )
```

# Assumption checking: Bootstrap means and t-statistics from trial data

TODO: Only done for large reward region and high control condition, pooled across training and test 

```{r}
set.seed(1115)
nsim <- 100000 # number of new datasets
boot_mean <- rep(0.0,nsim) # initialize a vector to store the means in the new datasets

trial_highcontrol_largereward <- trial_df %>%
  filter(
    reward == "Large Reward Region",
    control == "High Control"
  )

for(i in 1:nsim) { # iterate for each bootstrap sample
  y <- trial_highcontrol_largereward[sample(1:nrow(trial_highcontrol_largereward),size=nrow(trial_highcontrol_largereward),replace=TRUE),]$trial_score # sample data
  boot_mean[i] <- mean(y) # compute mean and store it in an element of boot_mean
}

ggplot(data.frame(mean=boot_mean), aes(x=mean)) + geom_histogram(binwidth=0.5) + 
  stat_function(fun = function(x) dnorm(x, mean = mean(boot_mean), sd = sd(boot_mean)) * 0.5 * nsim, linetype=2) + ylab("count")
```

```{r}
set.seed(2231)
nsim <- 100000 # number of new datasets
boot_t <- rep(0.0,nsim)
s_mean <- mean(trial_highcontrol_largereward$trial_score)
y <- matrix(trial_highcontrol_largereward[sample(1:nrow(trial_highcontrol_largereward),size=nsim*nrow(trial_highcontrol_largereward),replace=TRUE),]$trial_score,ncol=nsim)
boot_t <- apply(y, MARGIN=2, FUN=function(x) t.test(x, mu=s_mean)$statistic)

ggplot(data.frame(t=boot_t), aes(x=t)) + geom_histogram(binwidth=.1) +
  stat_function(fun = function(x) dt(x, df = nrow(trial_highcontrol_largereward) - 1) * .1 * nsim, linetype=2) + ylab("count")
```


```{r}
t_test_results2 <- t_test_results %>%
  select(-c(9:12))
```




